lifespan_eda

### 2025-01-13 eda/lifespan_eda.Rmd
update raw data with corrected VETSA, correct age_days to account for gestation

Load data from Jakob - updated with euler numbers!
```{r}
cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)
```

drop UKB for sharing
```{r, eval=FALSE}
cn.df_dropUKB <- cn.df %>%
  dplyr::filter(study != "UKB")
fwrite(cn.df_dropUKB, file=paste0(base_path, "/data/lifespan_dataset_dkatlas_CNonly_euler_noukb.csv"))
```

@@

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
```

@@

```{r}
cn.df.noinf <- cn.df.unique %>%
  dplyr::filter(age_days >= (365.25*3)) %>%
  mutate(study = droplevels(study),
         site = droplevels(site))
nrow(cn.df.unique) - nrow(cn.df.noinf)
n_unique(cn.df.unique$study) - n_unique(cn.df.noinf$study)
min(cn.df.noinf$age_days)
```

@@

#age
min(cn.df.imp$age_days)/365.25
max(cn.df.imp$age_days)/365.25
mean(cn.df.imp$age_days)/365.25

@@

drop UKB for sharing
```{r}
cn.df.imp_dropUKB <- cn.df.imp %>%
  dplyr::filter(study != "UKB")
fwrite(cn.df.imp_dropUKB, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler_noukb.csv"))
```

>>>>

Load data from Jakob - updated with euler numbers and corrected VETSA study
```{r}
cn.df <- fread(paste0(base_path,"/data/v1_2_1_nature.csv"), stringsAsFactors = TRUE)
```

@@

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
nrow(cn.df) - nrow(cn.df.unique)
```

@@

```{r}
cn.df.noinf <- cn.df.unique %>%
  dplyr::filter(age_days >= (365.25*3+280)) %>%
  mutate(study = droplevels(study),
         site = droplevels(site))
nrow(cn.df.unique) - nrow(cn.df.noinf)
n_unique(cn.df.unique$study) - n_unique(cn.df.noinf$study)
min(cn.df.noinf$age_days)
```

@@

#age
min(cn.df.imp$age_days)/365.25-(280/365.25)
max(cn.df.imp$age_days)/365.25-(280/365.25)
mean(cn.df.imp$age_days)/365.25-(280/365.25)

@@
